Genetic influence on wood density variation in  white-spruce (*Picea glauca*)

### 2019-10-15 Article_AS.Rmd
Discussion

---
title: "On the eletrodynamic of moving bodies"
author: "André D. Soro, Patrick Lenz, Jean Bousquet, Mariana Hassegawa, Jean-Romain Roussel, Alexis Achim"
output: 
  pdf_document: default
  word_document: default
bibliography: biblio.bibtex
editor_options: 
  chunk_output_type: inline
---

@@

Wood cores were extracted to a thickness of 1,68 mm from the south facing side of each tree 15-year-old trees in 2015 with a breast-height diameter (DBH15, 1.3 m above ground). Wood cores were extracted using increment borer which is a specialized tool for extracting wood cores from a living tree with relatively minor damage to the plant itself. These cores were used to determine ring width and wood density. Density was determined by scanning incremental cores from pith-to-bark along the radial face at a resolution of 25 $\mu$m in the QTRS-01X ring analyzer (Quintek Measurement Systems Inc., Knoxville, TN, USA). A weighted mean density line was obtained by summing the product between the annual density measurements and the ratio of the respective radial areas to the total radial area. Wood cores were conditioned to about 7% moisture before scanning. Determination of the wood density makes it possible to calculate a quality index of it. From the density profiles, the limits of the annual growth rings were determined together with the relative proportions of earlywood and latewood in each annual ring. Ring width determined through information of densitometry profiles. The transition between earlywood and the latewood is determined using a script developed, to check for possible errors (missing rings, abnormally low or high readings) and calculate the transition point between the initial wood and the final wood. The script uses the average between the minimum density and the maximum density to obtain the transition density threshold ($D_t$).

@@

## Model Development
A model was developed to describe the patterns of variation in mean annual ring density in the stem in white spruce for the purpose of modeling the known density variation portion and subtracting it to better observe the rest of the density variations. We had hierarchical structure of the data, with observations on annual rings nested in the cores, cores nested in the trees, and trees nested in blocks. Unlike some previous studies [@franceschini2010; @gardiner2011] that used mixed models to describe the radial density profile, we adjusted the curve of our equation to each core for a better description of the profiles. We, therefore, used the optim function of the stats package in the R [@R2018] statistical programming environment to find the best parameters iteratively for each tree core. 
As in the study of @Xiang2014, pith-to-bark wood density profiles in our data were different among cores. A nonlinear equation was fit for each core to describe ring density profiles, producing different parameter estimates for each core. The radial pattern of wood density in white spruce can be described as follows:

@@

## Spearman’s rank correlation coefficient (Spearman’s rho) 
We used a Spearman’s rank correlation coefficient to make a comparison of the response of the tree families between the two study sites. The elements for which we calculated Spearman’s rho are: V for each of the tree families between two study sites and the three parameters to estimate from the model. The $cor.test$ function in the R [@R2018] statistical programming environment was used to calculate Spearman’s rho. We used a 95 percent confidence interval.

@@

## Quantitative genetic analyses
Variance components for genetic analysis were estimated using ASReml 3 [@Butler2018] based on linear mixed-effects model using as follows:

@@

When comparing the index V between study sites, there is greater variability in Asselin than in Saint-Casimir. A summary of index V values can be found in figure 8. The maximum value is 1155.8 in Asselin, and 746 for the Saint-Casimir site, which represents a huge difference. \

|Rank correlation  |Rho  |P-value |
|:-----------------|----:|:-------|
|Density residus   | 0.57|8.7^-09^|
|Index V           | 0.08|0.43    |
|$b_0$             | 0.70|2.2^-16^|
|$b_1$             | 0.80|2.2^-16^|
|$b_2$             | 0.60|2.2^-16^|
Table: Sperman’s Rang correlation

@@

The dataset showed a typical radial pattern of wood density already observed in spruce trees in other researchers [@Alteyrac2006; @GARDINER2011; @Xiang2014] with the density initially decreasing outwards from the pith before slowly increasing towards an asymptotic value. This pattern corresponds to the type II pith-to-bark profile described by @panshin1980. The high density of wood adjacent to the pith has been attributed to high earlywood density and low latewood proportion [@alteyrac2007; @koubaa2007]. In addition, in the first years of a tree's life, it produces "flexure wood" that is high density [@telewski1989]. This type of wood is formed because the low rigidity of the young trees that allows them to withstand environmental constraints without breaking. The proportion of this wood decreases rapidly in the first few rings. The increase in density that follows the decline of the latter is related to the proportion of latewood that increase in mature wood [@clark2007]. Similar hypotheses have been proposed to explain radial variation in wood properties by @Lachenbruch2011. However, this radial pattern does not take into account the interannual variability of density. \

The modeling of individual profiles is suitable for the description of the wood density of our type II data. The prediction of the model on all of the data that can be seen in Figures 3 and 4 is an example of the modeling that is done for individual profiles. The choice of modeling the variation of the density of each of the trees made it possible to adapt the model to the variations between the trees. The fact that the model equation is composed of two components has given sufficient flexibility to properly describe type II density. This type of equation has been used previously in the study of @Xiang2014, where the authors used a three-component equation. In their case, the third component was used to describe a decline near the bark, which is not the case with what is observed in our study.\

@@

Genetic parameter estimates were used to evaluate the variation of density as a selection trait for wood quality. With regard to heritability of density, it is high at both sites. This strong density heritability is similar to previous studies in white spruce [Nelson 1963 in @mothe1983; @Corriveau1991; @Yanchuk1993; @Ivkovich2002a; @Ivkovich2002b; @Beaulieu2006]. All of these studies reported values between 0.36 and 0.80. Average density of our data is therefore highly controlled by genetics as might be expected.
In contrast to density, annual ring width has low to moderate heritability in each site and also for combined sites. This result, which is in agreement with previous studies [@Corriveau1991; @Ivkovich2002a], is explained by the difference we observe when we take the sites separately and when they are combined can be explained by a genotype-environment interaction [@kremer1986], this happens when conditions from one environment to another do not have the same effect on the genotype. This result suggests that, if trees are to be selected for reforestation as a result of this study, the territory be divided into relatively homogeneous subunits, within which the recommended sources will be able to keep their superiority and can be used without risk of maladjustment [@beaulieu1989stabilite].
The parameters of the model, as for them, they have a low heritability either for each site separately or for the combined sites. The parameter $b_0$, has a higher heritability than the other parameters but remains low. This can be explained by the fact that this parameter is very little influenced by cambial age because it is described the density close to pith. In fact, $b_0$ is the ordinate at the origin and thus little influenced by the sites (environmental conditions). Since parameters $b_1$ and $b_2$ are the parameters that describe the change in density over the years of tree growth, they are more likely to be influenced by environmental conditions.
The index V has a heritability close to zero for each of the sites (Asselin 6.10^-6^ and Saint-Casimir 3.10^-6^) and the two sites combined (1.10^-7^). The low genetic control of the index V shows that in addition to the fact that the interannual density variability is controlled by environment [@becker1995; @huang2007; @Nicault2010; @zaehle2014], its amplitude is also controlled by environment. One hypothesis that can explain the low heritability of the V index is the young age of our trees. Indeed, the study of @Gaspar2008 showed an increase in the heritability of the components of density with age. As a result, we can hope for an increase in the heritability of the index V, which represents the amplitude of the interannual variation in wood density.

@@

This study assessed the variability of wood density in white spruce. Using the optimal function of the stats package to find the best parameters iteratively for each tree allowed to have an adjustment for the density profile of each of the trees in the study. Thus, the index of variability of the density was determined from the profile of each tree. The low heritability of this index puts a damper on its use for identifying potential tree family of interest for breeding programs from a low interannual variability perspective of wood density. As for the heritability of ring density and width, it is similar to what can be found in previous studies. The application of the method used in this study to older trees can be interesting in order to make a comparison and see if the heritability is the same and if this index can be used in this case to select families of trees for genetic improvement.

>>>>

---
title: "Genetic influence on wood density variation in  white-spruce (*Picea glauca*)"
author: André D. Soro, Patrick Lenz, Jean Bousquet, Mariana Hassegawa, Jean-Romain
  Roussel, Alexis Achim
editor_options:
  chunk_output_type: inline
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
bibliography: biblio.bibtex
---

@@

Wood density and rings width was measured from wood core taken from the sampled trees. Wood cores were extracted to a thickness of 1,68 mm from the south facing side of each tree 15-year-old trees in 2015 with a breast-height diameter (DBH15, 1.3 m above ground). Wood cores were extracted using increment borer which is a specialized tool for extracting wood cores from a living tree with relatively minor damage to the plant itself. These cores were used to determine ring width and wood density. Density was determined by scanning incremental cores from pith-to-bark along the radial face at a resolution of 25 $\mu$m in the QTRS-01X ring analyzer (Quintek Measurement Systems Inc., Knoxville, TN, USA). A weighted mean density line was obtained by summing the product between the annual density measurements and the ratio of the respective radial areas to the total radial area. Wood cores were conditioned to about 7% moisture before scanning. Determination of the wood density makes it possible to calculate a quality index of it. From the density profiles, the limits of the annual growth rings were determined together with the relative proportions of earlywood and latewood in each annual ring. Ring width determined through information of densitometry profiles. The transition between earlywood and the latewood is determined using a script developed, to check for possible errors (missing rings, abnormally low or high readings) and calculate the transition point between the initial wood and the final wood. The script uses the average between the minimum density and the maximum density to obtain the transition density threshold ($D_t$).

@@

## Model Development
A model was developed to describe the patterns of variation in mean annual ring density in the stem in white spruce for the purpose of modeling the known density variation portion and subtracting it to better observe the rest of the density variations. Our data has a hierarchical structure, with observations on annual rings nested in the cores, cores nested in the trees, and trees nested in blocks. Unlike some previous studies [@franceschini2010; @gardiner2011] that used mixed models to describe the radial density profile, we adjusted the curve of our equation to each core for a better description of the profiles. We, therefore, used the optim function of the stats package in the R [@R2018] statistical programming environment to find the best parameters iteratively for each tree core. 
As in the study of @Xiang2014, pith-to-bark wood density profiles in our data were different among cores. A nonlinear equation was fit for each core to describe ring density profiles, producing different parameter estimates for each core. The radial pattern of wood density in white spruce can be described as follows:

@@

## Spearman’s rank correlation coefficient (Spearman’s rho) 
We used a Spearman’s rank correlation coefficient to make a comparison of the response of the tree families between the two study sites. This comparison will allow determine the interaction between the environment and the family effect. The elements for which we calculated Spearman’s rho are: V for each of the tree families between two study sites and the three parameters to estimate from the model. The $cor.test$ function in the R [@R2018] statistical programming environment was used to calculate Spearman’s rho. We used a 95 percent confidence interval.

@@

## Quantitative genetic analyses
Analyses of variance and covariance for various wood density and tree growth traits studied were performed to estimate genetic parameters. Variance components for genetic analysis were estimated using ASReml 3 [@Butler2018] based on linear mixed-effects model using as follows:

@@

When comparing the index V between study sites, there is greater variability in Asselin than in Saint-Casimir. A summary of index V values can be found in figure 8. The mean value is 417.5 in Asselin, and 339.8 for the Saint-Casimir site, which represents a huge difference. \

|Rank correlation  |Rho  |P-value |
|:-----------------|----:|-------:|
|Density residus   | 0.57|8.7^-09^|
|Index V           | 0.08|0.43    |
|$b_0$             | 0.70|2.2^-16^|
|$b_1$             | 0.80|2.2^-16^|
|$b_2$             | 0.60|2.2^-16^|
Table: Sperman’s Rang correlation

@@

Our data showed a typical radial pattern of wood density already observed in spruce trees in other researchers [@Alteyrac2006; @GARDINER2011; @Xiang2014] with the density initially decreasing outwards from the pith before slowly increasing towards an asymptotic value. This pattern corresponds to the type II pith-to-bark profile described by @panshin1980. The high density of wood adjacent to the pith has been attributed to high earlywood density and low latewood proportion [@alteyrac2007; @koubaa2007]. Indeed, the high density of earlywood in the first years of growth is explained by the fact that in the first years of a tree's life, it produces "flexure wood" that is high density [@telewski1989]. This type of wood is formed because the low rigidity of the young trees that allows them to withstand environmental constraints without breaking. The proportion of this wood decreases rapidly in the first few rings. The increase in density that follows the decline of the latter is related to the proportion of latewood that increase in mature wood [@clark2007]. Similar hypotheses have been proposed to explain radial variation in wood properties by @Lachenbruch2011. However, this radial pattern does not take into account the interannual variability of density. \

The modeling of individual profiles is suitable for the description of the wood density of our type II data. The choice of an equation containing two separate components allowed enough flexibility to represent the large variation in density profiles. The first term in the equation is capable of describing the type II density pattern previously described, whilst the second term was included when it was necessary to model a density increase close to the bark. The prediction of the model on all data that can be seen in Figures 3 and 4 is an example of the modeling that is done for individual profiles. The relatively poor fit of the full model in this study could be explained by the fact that, despite its classification as a type II species, wood density profiles in white spruce are, in reality, extremely variable. The choice of modeling the variation of the density of each of the trees made it possible to adapt the model to the variations between the trees. The fact that the model equation is composed of two components has given sufficient flexibility to properly describe type II density. This type of equation has been used previously in the study of @Xiang2014, where the authors used a three-component equation. In their case, the third component was used to describe a decline near the bark, which is not the case with what is observed in our study.\

@@

Genetic parameter estimates were used to evaluate the variation of density as a selection trait for wood quality. With regard to heritability of density, it is high at both sites. This strong density heritability is similar to previous studies in white spruce [Nelson 1963 in @mothe1983; @Corriveau1991; @Yanchuk1993; @Ivkovich2002a; @Ivkovich2002b; @Beaulieu2006]. All of these studies reported values between 0.36 and 0.80. Average density of our data is therefore highly controlled by genetics as might be expected.
In contrast to density, annual ring width has low to moderate heritability in each site and also for combined sites. This result, which is in agreement with previous studies [@Corriveau1991; @Ivkovich2002a], is explained by the difference we observe when we take the sites separately and when they are combined can be explained by a genotype-environment interaction [@kremer1986], this happens when conditions from one environment to another do not have the same effect on the genotype. This result suggests that, if trees are to be selected for reforestation as a result of this study, the territory be divided into relatively homogeneous subunits, within which the recommended sources will be able to keep their superiority and can be used without risk of maladjustment [@beaulieu1989stabilite].
The parameters of the model, as for them, they have a low heritability either for each site separately or for the combined sites. The parameter $b_0$, has a higher heritability than the other parameters but remains low. This can be explained by the fact that this parameter is very little influenced by cambial age because it is described the density close to pith. In fact, $b_0$ is the ordinate at the origin and thus little influenced by the sites (environmental conditions). Since parameters $b_1$ and $b_2$ are the parameters that describe the change in density over the years of tree growth, they are more likely to be influenced by environmental conditions.
The index V has a heritability close to zero for each of the sites (Asselin 6.10^-6^ and Saint-Casimir 3.10^-6^) and the two sites combined (1.10^-7^). The low genetic control of the index V shows that in addition to the fact that the interannual density variability is controlled by environment [@becker1995; @huang2007; @Nicault2010; @zaehle2014], its amplitude is also controlled by environment. One hypothesis that can explain the low heritability of the V index is the young age of our trees. Indeed, the study of @Gaspar2008 showed an increase in the heritability of the components of density with age. As a result, we can hope for an increase in the heritability of the index V for much older trees. \

@@

This study assessed the variability of wood density in white spruce. Using the optimal function of the stats package to find the best parameters iteratively for each tree allowed to have an adjustment for the density profile of each of the trees in the study. Thus, the index of variability of the density was determined from the profile of each tree. The low heritability of this index puts a damper on its use for identifying potential tree family of interest for breeding programs from a low interannual variability perspective of wood density. As for the heritability of ring density and width, it is similar to what can be found in previous studies. This study showed that the amplitude of the high frequency signal is variable between trees but that this variation is not controlled genetically. The application of the method used in this study to older trees can be interesting in order to make a comparison and see if the heritability is the same and if this index can be used in this case to select families of trees for genetic improvement.
